---
title: "Analysis with taxa concatenated to family level"
author: "segan"
date: '2022-11-28'
output:
  github_document:
    toc: true
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

# Load libraries

```{r results='hide'}
# libraries
pkgs <- c(
  "tidyverse",
  "phyloseq",
  "ggpubr",
  "ggplot2",
  "vegan",
  "reshape2",
  "plotly",
  "ampvis2",
  "microbiomeutilities",
  "RColorBrewer",
  "viridis",
  "biomeUtils",
  "microViz",
  "microbiome",
  "Rtsne",
  "umap",
  "yingtools2", #remotes::install_github("ying14/yingtools2")
  "scales",
  "ggrepel"
)

lapply(pkgs, require, character.only = TRUE)
theme_set(theme_bw())
```

#Create output folders
```{r}
dir.create(path = "./figures/heatmaps/all", recursive = TRUE)
dir.create(path = "./figures/heatmaps/heatmaps_by_study/phylum_level", recursive = TRUE)
dir.create(path = "./figures/heatmaps/heatmaps_by_study/family_level", recursive = TRUE)
dir.create(path = "./tables")
dir.create(path = "./figures/boxplot")
dir.create(path = "./figures/pie_charts")
dir.create(path = "./figures/alphadiv")
```


# Merge phyloseq files

```{r}
# Path to datasets drive
path_list <- list.files("./RData/primary_id", pattern=".Rda", full.names=TRUE)

# as load in each ps object rename to the project number
for (i in path_list)
{
  load(i)
  gdata::mv(from = "ps", to = gsub(".Rda", "", (sapply(strsplit(i, "/"), function(x)
    paste(x[4], sep = "/")))))
}

#Merge
ps_merge <-
  merge_phyloseq(
    ps_ERP021273,
    ps_ERP107582,
    ps_ERP108433,
    ps_ERP114897,
    ps_ERP120273,
    ps_ERP124577,
    ps_SRP187837,
    ps_SRP241247,
    ps_SRP247849,
    ps_SRP266568,
    ps_SRP279236,
    ps_SRP290868,
    ps_SRP307470,
    ps_SRP309452,
    ps_SRP320497,
    ps_SRP359931,
    ps_SRP369790,
    ps_SRP370478,
    ps_SRP383204
  )
```

Clean up taxonomy names 

```{r}
# taxonomy table
taxtab <- as.data.frame(ps_merge@tax_table)
#taxtab <- rownames_to_column(taxtab, var = "ASV_id")
# otu count table
otutab <- as.data.frame(ps_merge@otu_table)
#otutab <- rownames_to_column(otutab, var = "ASV_id")
# sample data table
samtab <- as.data.frame(ps_merge@sam_data)
# ref sequences
refseq <- refseq(ps_merge, FALSE)

sort(unique(taxtab$Phylum))
taxtab <-
  taxtab %>% 
  mutate(Phylum = ifelse(Phylum %in% "Bacillota", "Firmicutes [Bacillota]", Phylum)) %>% 
  mutate(Phylum = ifelse(Phylum %in% "Firmicutes", "Firmicutes [Bacillota]", Phylum)) %>% 
  mutate(Phylum = ifelse(Phylum %in% "Actinomycetota", "Actinobacteria [Actinomycetota]", Phylum)) %>% 
  mutate(Phylum = ifelse(Phylum %in% "Actinobacteria", "Actinobacteria [Actinomycetota]", Phylum)) %>% 
  mutate(Phylum = ifelse(Phylum %in% "Pseudomonadota", "Proteobacteria [Pseudomonadota]", Phylum))  %>% 
  mutate(Phylum = ifelse(Phylum %in% "Proteobacteria", "Proteobacteria [Pseudomonadota]", Phylum)) %>% 
  mutate(Class = ifelse(Class %in% "Actinomycetes", "Actinomycetia", Class)) %>% 
  mutate(Family = ifelse(Family %in% "Bacillales Family X. Incertae Sedis", "Bacillales X. (i.s.)", Family)) %>% 
  mutate(Family = ifelse(Family %in% "Clostridiales Family XVII. Incertae Sedis", "Clostridiales XVII. (i.s.)", Family)) %>% 
  mutate(Family = ifelse(Family %in% "Eubacteriales Family XII. Incertae Sedis", "Eubacteriales XII. (i.s.)", Family)) %>% 
  mutate(Family = ifelse(Family %in% "Eubacteriales Family XIII. Incertae Sedis", "Eubacteriales XIII. (i.s.)", Family)) %>% 
  mutate(Family = ifelse(Family %in% "Thermoanaerobacterales Family III. Incertae Sedis", "Thermoanaerobacterales III. (i.s.)", Family))

sort(unique(taxtab$Phylum))

taxmat <- as.matrix(taxtab)
otumat <- as.matrix(otutab)

OTU <- phyloseq::otu_table(otumat, taxa_are_rows = FALSE)
TAX <- phyloseq::tax_table(taxmat)
ps_merge <- phyloseq(OTU, TAX, samtab, refseq)


#Save phyloseq object as .Rda file
save(ps_merge, file = "./RData/ps_merge.Rda")
```

```{r}
load("./RData/ps_merge.Rda")

# ps_merge
# # phyloseq-class experiment-level object
# # otu_table()   OTU Table:         [ 83401 taxa and 1752 samples ]
# # sample_data() Sample Data:       [ 1752 samples by 106 sample variables ]
# # tax_table()   Taxonomy Table:    [ 83401 taxa by 7 taxonomic ranks ]
# # refseq()      DNAStringSet:      [ 83401 reference sequences ]

# # number of samples per project
# table(ps_merge@sam_data$bioreactor_model)
# # batch_Bas-Bellver_et_al_2020      7
# # batch_Korth_et_al_2022            420
# # batch_Kim_et_al_2022              115
# # batch_Kleigrewe_et_al_2022        27
# # batch_Elmen_et_al_2020            84
# # ARCOL_Leclerc_et_al_2021          46
# # M-ARCOL_Deschamps_et_al_2020      121
# # SHIME_based_Gaisawat_2020         96
# # SHIME_Firrman_et_al_2021          50
# # SHIME_Liu_et_al_2022              30
# # SHIME_Ma_et_al_2022               45
# # M-SHIME_Chassaing_et_al_2017      151
# # M-SHIME_Abbeele_2021              75
# # SIMGI_Zorraquin_et_al_2021        6
# # single_stage_Xu_et_al_ 2019       108
# # TIM-2_Larsen_et_al_2019           132
# # TIM-2_Vieira_et_al_2021           60
# # Infors_Oliphant_et_al_2019        62
# # PolyFermS_Pham_et_al_2019         117

```

# Concatenate taxonomy to family level 

Melt phyloseq object by Family level taxonomy

```{r}
grp_abund <- get_group_abundances(ps_merge,
                                  level = "Family",
                                  group = "sampleID",
                                  transform = "counts")
sort(unique(grp_abund$OTUID))
grp_abund <- dplyr::select(grp_abund, -sd_abundance)
count_family <- pivot_wider(grp_abund, names_from = "OTUID",
                    values_from = "mean_abundance")

# making OTU count table
OTU_mat <- count_family
otumat <- column_to_rownames(OTU_mat, var = "sampleID")
otumat <- as.matrix(otumat)

# making taxonomy table
TAX_names <- as.data.frame(colnames(count_family))
TAX_names$Family <- TAX_names$`colnames(count_family)`
colnames(TAX_names)[1] <- 'OTUID'
TAX_names <- TAX_names[-1,]
TAX_names$Genus <- ""
TAX_names$Species <- ""
taxdf <- as.data.frame(ps_merge@tax_table)
taxdf <- taxdf[,1:5]
taxdf <- taxdf[!duplicated(taxdf$Family), ]
TAX_mat <- left_join(TAX_names, taxdf, by = "Family")
TAX_mat <- dplyr::select(TAX_mat, OTUID, Kingdom, Phylum, Class, Order, Family, Genus, Species)
TAX_mat <- column_to_rownames(TAX_mat, var = "OTUID")
taxmat <- as.matrix(TAX_mat)
colnames(taxmat) <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")

# making sample data
sample_data <- as.data.frame(ps_merge@sam_data)
# make sure order matches the OTU count data.
idx<-match(OTU_mat$sampleID, sample_data$sampleID)
sample_data <- sample_data[idx,]
```

Re-make phyloseq object
```{r}
OTU <- phyloseq::otu_table(otumat, taxa_are_rows = FALSE)
TAX <- phyloseq::tax_table(taxmat)
ps_merge_fam <- phyloseq(OTU, TAX)
ps_merge_fam <- merge_phyloseq(ps_merge_fam, sample_data)
# saved merge concatenated family level taxa, all samples from all projects(inc. zero sum samples) 
save(ps_merge_fam, file = "./RData/ps_merge_fam.Rda")

# ps_merge_fam
# # phyloseq-class experiment-level object
# # otu_table()   OTU Table:         [ 248 taxa and 1752 samples ]
# # sample_data() Sample Data:       [ 1752 samples by 106 sample variables ]
# # tax_table()   Taxonomy Table:    [ 248 taxa by 7 taxonomic ranks ]
```

# Load data

Loading in merged `phyloseq` object concatenated at family level

```{r}
load("./RData/ps_merge_fam.Rda")

# ps_merge_fam
# # phyloseq-class experiment-level object
# # otu_table()   OTU Table:         [ 248 taxa and 1752 samples ]
# # sample_data() Sample Data:       [ 1752 samples by 106 sample variables ]
# # tax_table()   Taxonomy Table:    [ 248 taxa by 7 taxonomic ranks ]
```

## Excluding projects/samples

### 1. Exclude samples with 0 reads

Subset phyloseq object to exclude samples with 0 count

```{r}
#Identify which samples had sum of 0 reads
otumat2 <- as.data.frame(ps_merge_fam@otu_table)
otumat2 <- otumat2 %>% 
   mutate(sumrange = rowSums(across(where(is.numeric))))
idx<-which(otumat2$sumrange == 0)
otumat2 <- otumat2[idx,]
exclude <- rownames(otumat2)
exclude

# remove these from phyloseq object
ps_merge_fam <- subset_samples(ps_merge_fam, !sampleID %in%
                        exclude)
# ps_merge_fam
# # phyloseq-class experiment-level object
# # otu_table()   OTU Table:         [ 248 taxa and 1689 samples ]
# # sample_data() Sample Data:       [ 1689 samples by 106 sample variables ]
# # tax_table()   Taxonomy Table:    [ 248 taxa by 7 taxonomic ranks ]
```

### 2. Exclude samples coming from faeces

```{r}
ps_merge_fam <- subset_samples(ps_merge_fam, include_sample == "yes")

# ps_merge_fam
# # phyloseq-class experiment-level object
# # otu_table()   OTU Table:         [ 248 taxa and 1619 samples ]
# # sample_data() Sample Data:       [ 1619 samples by 106 sample variables ]
# # tax_table()   Taxonomy Table:    [ 248 taxa by 7 taxonomic ranks ]
table(ps_merge_fam@sam_data$bioreactor_model)
```

## Set order

Manipulating `ps_merge_fam@sam_data` variables

```{r}
ps_merge_fam@sam_data$bioreactor_type <-
  gsub(".*^M-", "", sapply(strsplit(
    ps_merge_fam@sam_data$bioreactor_model, "_"
  ), "[", 1))

ps_merge_fam@sam_data$stage <-
  as.character(ps_merge_fam@sam_data$stage)

ps_merge_fam@sam_data$ndonor <-
  factor(sapply(strsplit(ps_merge_fam@sam_data$ndonor_type, "_"), "[", 1),
    levels = c("1", "2", "3", "4", "5", "6", "7", "8", "12"))

ps_merge_fam@sam_data$donor_type <-
  sapply(strsplit(ps_merge_fam@sam_data$ndonor_type, "_"), "[", 2)
```

Set order for variables
```{r}
#Set order for `bioreactor_model`
ps_merge_fam@sam_data$bioreactor_model = factor(ps_merge_fam@sam_data$bioreactor_model, levels = c(
  "batch_Bas-Bellver_et_al_2020",
  "batch_Elmen_et_al_2020",
  "batch_Kim_et_al_2022",
  "batch_Kleigrewe_et_al_2022",
  "batch_Korth_et_al_2022",
  "single_stage_Xu_et_al_ 2019",
  "SHIME_based_Gaisawat_2020",
  "SHIME_Firrman_et_al_2021",
  "SHIME_Liu_et_al_2022",
  "SHIME_Ma_et_al_2022",
  "M-SHIME_Abbeele_2021",
  "M-SHIME_Chassaing_et_al_2017",
  "SIMGI_Zorraquin_et_al_2021",
  "TIM-2_Larsen_et_al_2019",
  "TIM-2_Vieira_et_al_2021",
  "Infors_Reese_et_al_2018",
  "PolyFermS_Pham_et_al_2019",
  "ARCOL_Leclerc_et_al_2021",
  "M-ARCOL_Deschamps_et_al_2020"
  ))

#Set order for `bioreactor_type`
ps_merge_fam@sam_data$bioreactor_type = factor(
  ps_merge_fam@sam_data$bioreactor_type,
  levels = c(
    "batch",
    "single",
    "SHIME",
    "SIMGI",
    "TIM-2",
    "Infors",
    "PolyFermS",
    "ARCOL"
  )
)
```

## Set colours

```{r}
cols_model <- c(
  "#a6bddb",
  "#74a9cf",
  "#3690c0",
  "#0570b0",
  "#034e7b",
  "#4d4d4d",
  "#d4b9da",
  "#c994c7",
  "#df65b0",
  "#ce1256",
  "#980043",
  "#67001f",
  "#252525",
  "#f46d43",
  "#fdae61",
  "#fee090",
  "#d9ef8b",
  "#41ab5d",
  "#005a32"
)

cols_type <- c(
  "#0570b0",
  "#4d4d4d",
  "#ce1256",
  "#252525",
  "#f46d43",
  "#fee090",
  "#d9ef8b",
  "#005a32"
)

shape_type = c(0, 1, 2, 3, 4, 5, 6, 7)
```

## Transform to compositional data

```{r}
ps_merge_fam_comp = transform_sample_counts(ps_merge_fam, function(x) sqrt(x/sum(x)))

##Can filter to min abundance??
#ps_merge_fam2_filt = filter_taxa(ps_merge_fam2, function(x) var(x) > 1e-05, TRUE)
```

## Make ampvis2 object

```{r}
# require the devtools package to source gist
if (!require("devtools"))
  install.packages("devtools")

#source the phyloseq_to_ampvis2() function from the gist
devtools::source_gist("8d0ca4206a66be7ff6d76fc4ab8e66c6")

# convert to ampvis2 and then save
ampvis2_fam <- phyloseq_to_ampvis2(ps_merge_fam)
ampvis2_fam_comp <- phyloseq_to_ampvis2(ps_merge_fam_comp)
```

```{r}
rm(count_family, grp_abund, otumat2)
rm(OTU_mat, otumat, sample_data, TAX_mat, TAX_names, taxdf, taxmat)
rm(OTU, TAX, exclude, idx)
```


# Library size of samples

```{r}
n <- n_distinct(ps_merge_fam@sam_data$primary_id)
df <-
  as.data.frame(sample_data(ps_merge_fam)) # Make ggplot-friendly data.frame
df$LibrarySize <- sample_sums(ps_merge_fam)
df <- df[order(df$LibrarySize), ]
df$Index <- seq(nrow(df))

df$bioreactor_model = factor(df$bioreactor_model, levels = c(
  "batch_Bas-Bellver_et_al_2020",
  "batch_Elmen_et_al_2020",
  "batch_Kim_et_al_2022",
  "batch_Kleigrewe_et_al_2022",
  "batch_Korth_et_al_2022",
  "single_stage_Xu_et_al_ 2019",
  "SHIME_based_Gaisawat_2020",
  "SHIME_Firrman_et_al_2021",
  "SHIME_Liu_et_al_2022",
  "SHIME_Ma_et_al_2022",
  "M-SHIME_Abbeele_2021",
  "M-SHIME_Chassaing_et_al_2017",
  "SIMGI_Zorraquin_et_al_2021",
  "TIM-2_Larsen_et_al_2019",
  "TIM-2_Vieira_et_al_2021",
  "Infors_Reese_et_al_2018",
  "PolyFermS_Pham_et_al_2019",
  "ARCOL_Leclerc_et_al_2021",
  "M-ARCOL_Deschamps_et_al_2020"
  ))

libQC <-
  ggplot(data = df,
         aes(
           x = bioreactor_model,
           y = LibrarySize,
           color = bioreactor_model
         )) + geom_point() +
  scale_color_manual(values = cols_model) +
  xlab("Bioreactor model") + ylab("Library size") +
  theme(axis.text.x = element_text(angle = 270, hjust = 0, vjust = 0, size =12), 
    axis.text.y = element_text(size =12),
    legend.position = "none")

ggsave(
  filename = "library_size.png",
  plot = libQC,
  path = "./figures",
  width = 10,
  height = 10,
  dpi = 300)
```

## Check for samples with low reads

```{r}
data.class(df$LibrarySize)
# summary of number of samples per project
summary <- df %>%
    group_by(primary_id) %>%
    summarise(count = n())
summary
# check with cut off at 5,000 reads
lowreads <- subset(df, LibrarySize < 5000)
tab1 <- lowreads %>%
    group_by(primary_id) %>%
    summarise(count = n())
tab1
```

# Heat maps

## All projects by taxa

```{r}
heatPhylum <- amp_heatmap(ampvis2_fam,
            group_by = "bioreactor_model",
            tax_aggregate = "Phylum",
            tax_show = 10,
            color_vector = c("white", 
                             "#e5ba52", 
                             "#ab7ca3", 
                             "#9d02d7", 
                             "#0030bf"),
            plot_colorscale = "sqrt",
            plot_values = TRUE,
            showRemainingTaxa = TRUE,
            round = 1) +
  theme(axis.text.x = element_text(angle = 45, size=10, vjust = 1),
        axis.text.y = element_text(size=10, face = "italic"),
        legend.position="right")

ggsave(
  "heatmap_Phylum.png",
  heatPhylum,
  path = "./figures/heatmaps/all",
  width = 25,
  height = 25,
  units = "cm"
)

heatFamily <- amp_heatmap(ampvis2_fam,
            group_by = "bioreactor_model",
            tax_aggregate = "Family",
            tax_show = 40,
            color_vector = c("white", 
                             "#e5ba52", 
                             "#ab7ca3", 
                             "#9d02d7", 
                             "#0030bf"),
            plot_colorscale = "sqrt",
            plot_values = TRUE,
            showRemainingTaxa = TRUE,
            round = 1) +
  theme(axis.text.x = element_text(angle = 45, size=10, vjust = 1),
        axis.text.y = element_text(size=10, face = "italic"),
        legend.position="right")

ggsave(
  "heatmap_Family.png",
  heatFamily,
  path = "./figures/heatmaps/all",
  width = 25,
  height = 25,
  units = "cm"
)
```

## By study

Producing heatmaps by study

### Phylum level

```{r heatmapPrjPhylum}
loop.vector <- unique(ampvis2_fam$metadata$bioreactor_model)

for (i in loop.vector){
  
  amp_sub <- amp_subset_samples(ampvis2_fam, 
                                 bioreactor_model %in% i)
  
  heatmap <- amp_heatmap(amp_sub,
            tax_aggregate = "Phylum",
            tax_show = 20,
            color_vector = c("white", 
                             "#e5ba52", 
                             "#ab7ca3", 
                             "#9d02d7", 
                             "#0030bf"),
            plot_colorscale = "sqrt",
            plot_values = TRUE,
            showRemainingTaxa = TRUE) +
  theme(axis.text.x = element_text(angle = 45, size=10, vjust = 1),
        axis.text.y = element_text(size=12),
        legend.position="right") +
  ggtitle(i)
  
  ggsave(
    filename = paste0(
      "./figures/heatmaps/heatmaps_by_study/phylum_level/",
      i,
      ".pdf"
    ),
    heatmap,
    width = 40,
    height = 25,
    units = "cm"
  )
}
```

### Family level

```{r heatmapPrjFamily}
loop.vector <- unique(ampvis2_fam$metadata$bioreactor_model)

for (i in loop.vector){
  
  amp_sub <- amp_subset_samples(ampvis2_fam, 
                                 bioreactor_model %in% i)
  
  heatmap <- amp_heatmap(amp_sub,
            tax_aggregate = "Family",
            tax_show = 20,
            color_vector = c("white", 
                             "#e5ba52", 
                             "#ab7ca3", 
                             "#9d02d7", 
                             "#0030bf"),
            plot_colorscale = "sqrt",
            plot_values = TRUE,
            showRemainingTaxa = TRUE) +
  theme(axis.text.x = element_text(angle = 45, size=10, vjust = 1),
        axis.text.y = element_text(size=12),
        legend.position="right") +
  ggtitle(i)
  
  ggsave(
    filename = paste0(
      "./figures/heatmaps/heatmaps_by_study/family_level/",
      i,
      ".pdf"
    ),
    heatmap,
    width = 40,
    height = 25,
    units = "cm"
  )
}
```

# Ordination - non linear methods
## t-SNE plots with compositional object

```{r}
set.seed(2)
dst <- phyloseq::distance(ps_merge_fam_comp, method = "euclidean")
tsne <- Rtsne(dst, is_distance = TRUE)
tsnedata <- data.frame(tsne$Y) 
colnames(tsnedata)[1] <- "tSNE1"
colnames(tsnedata)[2] <- "tSNE2"
tsnedata$bioreactor_model <- ps_merge_fam_comp@sam_data$bioreactor_model
tsnedata$region_16S <- ps_merge_fam@sam_data$region_16S
tsnedata$vitamins <- ps_merge_fam@sam_data$vitamins
tsnedata$country_donor <- ps_merge_fam@sam_data$country_donor
tsnedata$CNratio <- ps_merge_fam@sam_data$CNratio
tsnedata$stage <- ps_merge_fam@sam_data$stage
tsnedata$stages <- ps_merge_fam@sam_data$stages
tsnedata$media_based_on <- ps_merge_fam@sam_data$media_based_on
tsnedata$mucin <- ps_merge_fam@sam_data$mucin
tsnedata$predigestion <- ps_merge_fam@sam_data$predigestion
tsnedata$ndonor_type <- ps_merge_fam@sam_data$ndonor_type
tsnedata$bioreactor_type <- ps_merge_fam@sam_data$bioreactor_type

tsne_plot <- ggplot(tsnedata, aes(x = tSNE1, y = tSNE2, color = bioreactor_model)) +
  geom_point() +
  labs(color = "Bioreactor model", shape = "Bioreactor type") +
  theme(aspect.ratio = 1, text = element_text(size = 13)) +  
  scale_color_manual(values = cols_model)

ggsave(
  "distance_comp_tsne_hellinger.png",
  tsne_plot,
  path = "./figures/tSNE",
  width = 25,
  height = 20,
  units = "cm"
)

# # Interactive view to understand patterns
# plotly::ggplotly(tsne_plot)

## Facet by type to be able to visualize better
tsne_plot2 <- ggplot(tsnedata, aes(x = tSNE1, y = tSNE2, color = bioreactor_model)) +
  geom_point() +
  labs(color = "Bioreactor model") +
  theme(aspect.ratio = 1) +  
  scale_color_manual(values = cols_model) +
  facet_wrap(~bioreactor_type) +
  theme(legend.position="bottom") + guides(col = guide_legend(ncol = 4)) +
  theme(legend.title=element_blank())

ggsave(
  "distance_comp_tsne_hellinger_facet.png",
  tsne_plot2,
  path = "./figures/tSNE",
  width = 25,
  height = 20,
  units = "cm"
)
```

#### Effect of additional parameters on community structure

Supplementary figure 

```{r}
tsne_plot_16S <- ggplot(tsnedata, aes(x = tSNE1, y = tSNE2, color = region_16S)) +
  geom_point() +
  labs(color = "16S rRNA region") +
  theme(aspect.ratio = 1) +  
  scale_color_viridis(discrete=TRUE)

tsne_plot_CNratio <- ggplot(tsnedata, aes(x = tSNE1, y = tSNE2, color = as.character(CNratio))) +
  geom_point() +
  labs(color = "C:N ratio") +
  theme(aspect.ratio = 1) +
  scale_color_viridis(discrete=TRUE)

tsne_plot_vitamins <- ggplot(tsnedata, aes(x = tSNE1, y = tSNE2, color = vitamins)) +
  geom_point() +
  labs(color = "vitamins") +
  theme(aspect.ratio = 1) +  
  scale_color_viridis(discrete=TRUE)

tsne_plot_country <- ggplot(tsnedata, aes(x = tSNE1, y = tSNE2, color = country_donor)) +
  geom_point() +
  labs(color = "Country") +
  theme(aspect.ratio = 1) +  
  scale_color_viridis(discrete=TRUE)

## Arrange and save plot for supplementary information
gg_arrange_1 <-
  ggarrange(
    tsne_plot_16S,
    tsne_plot_CNratio,
    tsne_plot_country,
    tsne_plot_vitamins,
    labels = c("A", "B", "C", "D")
  )

ggsave(
  "distance_comp_tsne_hellinger_supp1.png",
  gg_arrange_1,
  path = "./figures/tSNE",
  width = 25,
  height = 25,
  units = "cm"
)
```

Supplementary figure

```{r}
tsne_plot_stage <- ggplot(tsnedata, aes(x = tSNE1, y = tSNE2, color = as.character(stage))) +
  geom_point() +
  labs(color = "Stage") +
  theme(aspect.ratio = 1) +  
  scale_color_viridis(discrete=TRUE)

tsne_plot_gut <- ggplot(tsnedata, aes(x = tSNE1, y = tSNE2, color = stages)) +
  geom_point() +
  labs(color = "Gut compartment") +
  theme(aspect.ratio = 1) +  
  scale_color_viridis(discrete=TRUE)

tsne_plot_predigestion <- ggplot(tsnedata, aes(x = tSNE1, y = tSNE2, color = predigestion)) +
  geom_point() +
  labs(color = "Predigestion") +
  theme(aspect.ratio = 1) +  
  scale_color_viridis(discrete=TRUE)

tsne_plot_mucin <- ggplot(tsnedata, aes(x = tSNE1, y = tSNE2, color = as.character(mucin))) +
  geom_point() +
  labs(color = "Mucin") +
  theme(aspect.ratio = 1) +  
  scale_color_viridis(discrete=TRUE)

gg_arrange_2 <-
  ggarrange(
    tsne_plot_stage,
    tsne_plot_gut,
    tsne_plot_predigestion,
    tsne_plot_mucin,
    labels = c("A", "B", "C", "D")
  )

ggsave(
  "distance_comp_tsne_hellinger_supp2.png",
  gg_arrange_2,
  path = "./figures/tSNE",
  width = 25,
  height = 25,
  units = "cm"
)
```
## UMAP plots with compositional object
```{r}
# count
s <- get.samp(ps_merge_fam_comp)
otu <- get.otu(ps_merge_fam_comp, as.matrix = TRUE)
# depending on how otu table was original formatted may need to transpose
# rows = samples and columns = otus/taxa
otu <- t(otu)
umap <- umap(otu)
umapdata <- umap$layout %>%
  data.frame() %>%
  rownames_to_column("sample")
colnames(umapdata)[2] <- "UMAP1"
colnames(umapdata)[3] <- "UMAP2"
umapdata <- left_join(umapdata, s, by = "sample")
umap_plot <- ggplot(umapdata, aes(x = UMAP1, y = UMAP2, color = bioreactor_model)) +
  geom_point() +
  labs(color = "Bioreactor model", shape = "Bioreactor type") +
  theme(aspect.ratio = 1, text = element_text(size = 13)) +  
  scale_color_manual(values = cols_model)

ggsave(
  "distance_comp_umap_hellinger.png",
  umap_plot,
  path = "./figures/umap",
  width = 25,
  height =20,
  units = "cm",
  dpi = 300
)

## Facet by type to be able to visualize better
umap_plot2 <- ggplot(umapdata, aes(x = UMAP1, y = UMAP2, color = bioreactor_model)) +
  geom_point() +
  labs(color = "Bioreactor model") +
  theme(aspect.ratio = 1) +  
  scale_color_manual(values = cols_model) +
  facet_wrap(~bioreactor_type) +
  theme(legend.position="bottom") + guides(col = guide_legend(ncol = 4)) +
  theme(legend.title=element_blank())

ggsave(
  "distance_comp_umap_hellinger_facet.png",
  umap_plot2,
  path = "./figures/umap",
  width = 25,
  height = 20,
  units = "cm"
)
```

# Alpha diversity plots with statistical tests

Create simple alpha diversity plots with count data - with separate calculation and data.frame for statistical tests

## Count data - simple plot by bioreactor_model

```{r}
# group by `bioreactor_model`
alpha_div_model <-
  plot_richness(ps_merge_fam,
    measures = c("Observed", "Chao1", "Shannon", "InvSimpson"))
alpha_count_df <- alpha_div_model$data

p <-
  ggplot(alpha_count_df,
    aes(x = bioreactor_model, y = value, fill = bioreactor_model)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, fill = "white") +
  facet_wrap( ~ variable, scales = "free_y") +
  scale_color_manual(values = cols_model)  +
  theme(legend.position = "none",
    axis.text.x = element_text(angle = 90, size = 8)) + xlab("") + ylab("") 

ggsave(
  "count_bioreactor_model.png",
  p,
  path = "./figures/alphadiv",
  width = 25,
  height = 15,
  units = "cm"
)

### As there are too many pair-wise groups that are significant making a data.frame with statistical values to report in results text

## Observed
obs_results <- alpha_count_df %>% 
  filter(variable == "Observed") %>%
  statsExpressions::pairwise_comparisons(
    bioreactor_model,
    value,
    var.equal = FALSE,
    type = "non-parametric")
obs_results$variable <- "Observed"
obs_results <- obs_results %>%
  mutate(
    p.value.sig = case_when(
      p.value > 0.05 ~ "n.s.",
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*"
    )
  )

## Chao1
chao1_results <- alpha_count_df %>% 
  filter(variable == "Chao1") %>%
  statsExpressions::pairwise_comparisons(
    bioreactor_model,
    value,
    var.equal = FALSE,
    type = "non-parametric")
chao1_results$variable <- "Chao1"
chao1_results <- chao1_results %>%
  mutate(
    p.value.sig = case_when(
      p.value > 0.05 ~ "n.s.",
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*"
    )
  )

## Shannon
shan_results <- alpha_count_df %>% 
  filter(variable == "Shannon") %>%
  statsExpressions::pairwise_comparisons(
    bioreactor_model,
    value,
    var.equal = FALSE,
    type = "non-parametric")
shan_results$variable <- "Shannon"
shan_results <- shan_results %>%
  mutate(
    p.value.sig = case_when(
      p.value > 0.05 ~ "n.s.",
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*"
    )
  )

## Simpson
invsimp_results <- alpha_count_df %>% 
  filter(variable == "InvSimpson") %>%
  statsExpressions::pairwise_comparisons(
    bioreactor_model,
    value,
    var.equal = FALSE,
    type = "non-parametric")
invsimp_results$variable <- "InvSimpson"
invsimp_results <- invsimp_results %>%
  mutate(
    p.value.sig = case_when(
      p.value > 0.05 ~ "n.s.",
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*"
    )
  )

alpha_count_stats<- rbind(obs_results, chao1_results, shan_results, invsimp_results)
alpha_count_stats$expression <- as.character(alpha_count_stats$expression)
write.csv(alpha_count_stats, "./tables/alpha_count_stats.csv")

alpha_count_stats <- read.csv("./tables/alpha_count_stats.csv", header = TRUE)

alpha_count_stats_summary <- alpha_count_stats %>% 
  filter(p.value.sig == "*" | p.value.sig == "**" | p.value.sig == "***" & variable == "Observed") %>%
  group_by(group1, p.value.sig) %>% # Variable to be transformed
  count() %>% 
  group_by(group1) %>% 
  mutate(sum(n))
 
```

## Count data by `bioreactor_type` - with statistical analysis on plots

```{r}
p1 <- ggstatsplot::ggbetweenstats(
  data = dplyr::filter(alpha_count_df, variable == "Observed"),
  x = bioreactor_type,
  y = value,
  type = "non-parametric", #Kruskal-Wallis
  var.equal = FALSE, # Welch ANOVA
  plot.type = "box",
  results.subtitle = FALSE,
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = FALSE,
  bf.message = FALSE,
  # ggplot.component = list(theme(text = element_text(size = 15)))
) +
   scale_color_manual(values = cols_type) +
   theme(
     legend.position = "none",
     axis.title.x = element_blank()
   ) + ylab("Observed diversity")

p2 <- ggstatsplot::ggbetweenstats(
  data = dplyr::filter(alpha_count_df, variable == "Shannon"),
  x = bioreactor_type,
  y = value,
   type = "non-parametric", #Kruskal-Wallis
  var.equal = FALSE, # Welch ANOVA
  plot.type = "box",
  results.subtitle = FALSE,
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = FALSE,
  bf.message = FALSE,
) +
   scale_color_manual(values = cols_type) +
   theme(
     legend.position = "none",
     axis.title.x = element_blank()
   ) + ylab("Shannon diversity")

alpha_type <-
  ggarrange(
    p1,
    p2,
    nrow = 2
  )

ggsave(
  "count_bioreactor_type_stats.pdf",
  alpha_type,
  path = "./figures/alphadiv",
  width = 15,
  height = 25,
  units = "cm"
)
```

## Count data by donor variables - with statistical analysis on plots

```{r}
p1 <- ggstatsplot::ggbetweenstats(
  data = dplyr::filter(alpha_count_df, variable == "Shannon"),
  x = donor,
  y = value,
  type = "non-parametric",
  #Kruskal-Wallis
  var.equal = FALSE,
  # Welch ANOVA
  plot.type = "box",
  results.subtitle = TRUE,
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = FALSE,
  bf.message = FALSE
) +
  scale_color_manual(values = c(rep(
    "grey", n_distinct(alpha_count_df$donor)
  ))) +
  theme(legend.position = "none",
    axis.title.x = element_blank()) + ylab("Shannon diversity")

p2 <- ggstatsplot::ggbetweenstats(
  data = dplyr::filter(alpha_count_df, variable == "Shannon"),
  x = donor_type,
  y = value,
  type = "non-parametric",
  #Kruskal-Wallis
  var.equal = FALSE,
  # Welch ANOVA
  plot.type = "box",
  results.subtitle = TRUE,
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = FALSE,
  bf.message = FALSE
) +
  scale_color_manual(values = c(rep(
    "grey", n_distinct(alpha_count_df$donor_type)
  ))) +
  theme(legend.position = "none",
    axis.title.x = element_blank()) + ylab("Shannon diversity")

alpha_variables <-
  ggarrange(p1, p2,
    nrow = 2)

ggsave(
  "count_bioreactor_donor_stats.pdf",
  alpha_variables,
  path = "./figures/alphadiv/",
  width = 15,
  height = 25,
  units = "cm"
)
```

Supplementary information table with alpha diversity summary

```{r}
library(kableExtra)

df <- alpha_count_df %>%
  group_by(bioreactor_model, variable) %>%
  summarize(mean_value = mean(value), med = median(value), Sdev = sd(value))
df <- df[order(df$variable),]
df$mean_value <- round(df$mean_value, 2)
df$med <- round(df$med, 2)
df$Sdev <- round(df$Sdev, 2)
kbl(df, caption = "Group Rows") %>%
  kable_paper("striped", full_width = F) %>%
  pack_rows("Group 1", 1, 19) %>%
  pack_rows("Group 2", 20, 38) %>%
  pack_rows("Group 3", 39, 57) %>%
  pack_rows("Group 4", 58, 76)

df <- alpha_count_df %>%
  group_by(bioreactor_type, variable) %>%
  summarize(mean_value = mean(value), med = median(value), Sdev = sd(value))
df <- df[order(df$variable),]
df$mean_value <- round(df$mean_value, 2)
df$med <- round(df$med, 2)
df$Sdev <- round(df$Sdev, 2)
kbl(df, caption = "Group Rows") %>%
  kable_paper("striped", full_width = F) %>%
  pack_rows("Group 1", 1, 8) %>%
  pack_rows("Group 2", 9, 16) %>%
  pack_rows("Group 3", 17, 24) %>%
  pack_rows("Group 4", 25, 32)

df <- alpha_count_df %>%
  group_by(donor, variable) %>%
  summarize(mean_value = mean(value), med = median(value), Sdev = sd(value))
df <- df[order(df$variable),]
df$mean_value <- round(df$mean_value, 2)
df$med <- round(df$med, 2)
df$Sdev <- round(df$Sdev, 2)
kbl(df, caption = "Group Rows") %>%
  kable_paper("striped", full_width = F) %>%
  pack_rows("Group 1", 1, 5) %>%
  pack_rows("Group 2", 6, 10) %>%
  pack_rows("Group 3", 11, 15) %>%
  pack_rows("Group 4", 16, 20)

df <- alpha_count_df %>%
  group_by(donor_type, variable) %>%
  summarize(mean_value = mean(value), med = median(value), Sdev = sd(value))
df <- df[order(df$variable),]
df$mean_value <- round(df$mean_value, 2)
df$med <- round(df$med, 2)
df$Sdev <- round(df$Sdev, 2)
kbl(df, caption = "Group Rows") %>%
  kable_paper("striped", full_width = F) %>%
  pack_rows("Group 1", 1, 2) %>%
  pack_rows("Group 2", 3, 4) %>%
  pack_rows("Group 3", 5, 6) %>%
  pack_rows("Group 4", 7, 8)
```


# Rarefaction curves

Initial assessment of sequencing - one per sample with facet by the project

```{r}
rarecurve_prj <-
  amp_rarefaction_curve(ampvis2_fam, facet_by = "bioreactor_model",
    facet_scales = "free")


ggsave(
  filename = paste0("./figures/rarecurve.png"),
  rarecurve_prj,
  width = 40,
  height = 25,
  units = "cm"
)
```


Statistical assessment of sequencing depth for variable (primary_id) - one per primary_id (with confidence intervals)

```{r}
# set seed
set.seed(1)
# set subsample
subsamples <- seq(from = 0, to = 250000, by = 5000)
rarecurve <- plot_alpha_rcurve(
  ps_merge_fam,
  index = "observed",
  subsamples = subsamples,
  lower.conf = 0.025,
  upper.conf = 0.975,
  group = "primary_id",
  label.color = "brown3",
  label.size = 3,
  label.min = TRUE
)
# # change color of line
# mycols <- c("brown3", "steelblue")
# rarecurve_M <- rarecurve_M + scale_color_manual(values = mycols) +
#   scale_fill_manual(values = mycols)
```

# Alpha diversity - Hill numbers

```{r}
library(hilldiv)

otu_table <- data.frame(ps_merge_fam@otu_table)

otu_table <- t(otu_table)

hill_0 <- hill_div(otu_table, 0)
hill_1 <- hill_div(otu_table, 1)
hill_2 <- hill_div(otu_table, 2)
hill_3 <- hill_div(otu_table, 3)

hill_numbers <- data.frame(hill_0, hill_1, hill_2, hill_3)

sample_data <- data.frame(ps_merge_fam@sam_data)

sample_data <- data.frame(sample_data[, c(5, 10)], hill_numbers)

sample_data_long <-
  pivot_longer(sample_data,
    cols = c(hill_0, hill_1, hill_2, hill_3),
    names_to = "Hill_number")

colnames(sample_data_long)[1] <- "bioreactor_model"

sample_data_long$Hill_number  <-
  replace(sample_data_long$Hill_number,
    sample_data_long$Hill_number == "hill_0",
    0)
sample_data_long$Hill_number  <-
  replace(sample_data_long$Hill_number,
    sample_data_long$Hill_number == "hill_1",
    1)
sample_data_long$Hill_number  <-
  replace(sample_data_long$Hill_number,
    sample_data_long$Hill_number == "hill_2",
    2)
sample_data_long$Hill_number  <-
  replace(sample_data_long$Hill_number,
    sample_data_long$Hill_number == "hill_3",
    3)

sample_data_long$Hill_number <-
  as.numeric(sample_data_long$Hill_number)

## one figure
hill_numbers_plot_all <-
  sample_data_long %>%
  ggline(
    x = "Hill_number",
    y = "value",
    color = "bioreactor_model",
    add = c("mean_sd", "jitter"),
    facet.by = "bioreactor_type"
  ) +
  scale_color_manual(values = cols_model)

hill_numbers_plot_all <-
  ggpar(
    hill_numbers_plot_all,
    legend = "bottom",
    legend.title = "Model",
    xlab = "Diversity order (q)",
    ylab = "Diversity (Hill numbers)"
  )

ggsave(
  "hill_numbers_all_facet.svg",
  hill_numbers_plot_all,
  path = "./figures/alphadiv/",
  width = 35,
  height = 25,
  units = "cm"
)



hill_numbers_plot_all <-
  sample_data_long %>% ggline(x = "Hill_number",
    y = "value",
    color = "bioreactor_model",
    add = "mean_se")+
  scale_color_manual(values = cols_model)

hill_numbers_plot_all <-
  ggpar(
    hill_numbers_plot_all,
    legend = "top",
    legend.title = "Model",
    xlab = "Diversity order (q)",
    ylab = "Diversity (Hill numbers)",
    ylim = c(0, 40)
  )

ggsave(
  "hill_numbers_all.png",
  hill_numbers_plot_all,
  path = "./figures/alphadiv/",
  width = 35,
  height = 20,
  units = "cm"
)


## multi-panel
hill_numbers_plot_a <-
  sample_data_long %>% filter(
    bioreactor_model == "batch_Kim_et_al_2022" |
      bioreactor_model == "SHIME_Ma_et_al_2022" |
      bioreactor_model == "batch_Korth_et_al_2022" |
      bioreactor_model == "ARCOL_Leclerc_et_al_2021" |
      bioreactor_model == "SHIME_Liu_et_al_2022"
  ) %>% ggline(
    x = "Hill_number",
    y = "value",
    color = "bioreactor_model",
    add = c("mean_sd", "jitter")
  )

hill_numbers_plot_a <-
  ggpar(
    hill_numbers_plot_a,
    legend = "right",
    legend.title = "Model",
    xlab = "Diversity order (q)",
    ylab = "Diversity (Hill numbers)",
    ylim = c(0, 60)
  )

hill_numbers_plot_b <-
  sample_data_long %>% filter(
    bioreactor_model == "batch_Bas-Bellver_et_al_2020" |
      bioreactor_model == "batch_Kleigrewe_et_al_2022" |
      bioreactor_model == "M-SHIME_Chassaing_et_al_2017" |
      bioreactor_model == "M-SHIME_Abbeele_2021" |
      bioreactor_model == "TIM-2_Vieira_et_al_2021"
  ) %>% ggline(
    x = "Hill_number",
    y = "value",
    color = "bioreactor_model",
    add = c("mean_sd", "jitter")
  )

hill_numbers_plot_b <-
  ggpar(
    hill_numbers_plot_b,
    legend = "right",
    legend.title = "Model",
    xlab = "Diversity order (q)",
    ylab = "Diversity (Hill numbers)",
    ylim = c(0, 60)
  )

hill_numbers_plot_c <-
  sample_data_long %>% filter(
    bioreactor_model == "SHIME_based_Gaisawat_2020" |
      bioreactor_model == "SHIME_Firrman_et_al_2021" |
      bioreactor_model == "TIM-2_Larsen_et_al_2019" |
      bioreactor_model == "PolyFermS_Pham_et_al_2019" |
      bioreactor_model == "M-ARCOL_Deschamps_et_al_2020"
  ) %>% ggline(
    x = "Hill_number",
    y = "value",
    color = "bioreactor_model",
    add = c("mean_sd", "jitter")
  )

hill_numbers_plot_c <-
  ggpar(
    hill_numbers_plot_c,
    legend = "right",
    legend.title = "Model",
    xlab = "Diversity order (q)",
    ylab = "Diversity (Hill numbers)",
    ylim = c(0, 60)
  )

hill_numbers_plot_d <-
  sample_data_long %>% filter(
    bioreactor_model == "batch_Elmen_et_al_2020" |
      bioreactor_model == "single_stage_Xu_et_al_ 2019" |
      bioreactor_model == "SIMGI_Zorraquin_et_al_2021" |
      bioreactor_model == "Infors_Oliphant_et_al_2019"
  ) %>% ggline(
    x = "Hill_number",
    y = "value",
    color = "bioreactor_model",
    add = c("mean_sd", "jitter")
  )

hill_numbers_plot_d <-
  ggpar(
    hill_numbers_plot_d,
    legend = "right",
    legend.title = "Model",
    xlab = "Diversity order (q)",
    ylab = "Diversity (Hill numbers)",
    ylim = c(0, 60)
  )

multi_plot <-
  ggarrange(
    hill_numbers_plot_a,
    hill_numbers_plot_b,
    hill_numbers_plot_c,
    hill_numbers_plot_d
  )

ggsave(
  "Hill_numbers.pdf",
  multi_plot,
  path = "./figures/alphadiv/",
  width = 35,
  height = 20,
  units = "cm"
)

double_plot <-
  ggarrange(
    hill_numbers_plot_all,
    multi_plot,
    ncol = 1,
    nrow = 2,
    labels = c("A", "B"),
    common.legend = TRUE,
    font.label = list(size = 18)
  )

ggsave(
  "Hill_numbers_mix.pdf",
  double_plot,
  path = "./figures/alphadiv/",
  width = 35,
  height = 35,
  units = "cm"
)

ggsave(
  "Hill_numbers_mix.png",
  double_plot,
  path = "./figures/alphadiv/",
  width = 35,
  height = 35,
  units = "cm",
  dpi = 300
)
```

# Taxonomy comparison

```{r}
taxonomyBioreactors <- as.data.frame(ps_merge_fam@tax_table)

taxonomyFecal <- read.csv("./supporting_metadata/GMrepo_taxonomy.csv")

taxonomyFecal <-
  taxonomyFecal %>% 
  mutate(phylum = ifelse(phylum %in% "Bacillota", "Firmicutes [Bacillota]", phylum)) %>% 
  mutate(phylum = ifelse(phylum %in% "Firmicutes", "Firmicutes [Bacillota]", phylum)) %>% 
  mutate(phylum = ifelse(phylum %in% "Actinomycetota", "Actinobacteria [Actinomycetota]", phylum)) %>% 
  mutate(phylum = ifelse(phylum %in% "Actinobacteria", "Actinobacteria [Actinomycetota]", phylum)) %>% 
  mutate(phylum = ifelse(phylum %in% "Pseudomonadota", "Proteobacteria [Pseudomonadota]", phylum))  %>% 
  mutate(phylum = ifelse(phylum %in% "Proteobacteria", "Proteobacteria [Pseudomonadota]", phylum)) %>% 
  mutate(family = ifelse(family %in% "Bacillales family X. Incertae Sedis", "Bacillales X. (i.s.)", family)) %>% 
  mutate(family = ifelse(family %in% "Clostridiales family XVII. Incertae Sedis", "Clostridiales XVII. (i.s.)", family)) %>% 
  mutate(family = ifelse(family %in% "Eubacteriales family XII. Incertae Sedis", "Eubacteriales XII. (i.s.)", family)) %>% 
  mutate(family = ifelse(family %in% "Eubacteriales family XIII. Incertae Sedis", "Eubacteriales XIII. (i.s.)", family)) %>% 
  mutate(family = ifelse(family %in% "Thermoanaerobacterales family III. Incertae Sedis", "Thermoanaerobacterales III. (i.s.)", family))

bioreactor_vs_fecal <- data.frame(setdiff(taxonomyBioreactors$Family, taxonomyFecal$family))

bioreactor_vs_fecal <- data.frame(bioreactor_vs_fecal[-29,])

colnames(bioreactor_vs_fecal) <- "family_samples_only"

fecal_vs_bioreactor <- data.frame(setdiff(taxonomyFecal$family, taxonomyBioreactors$Family))

fecal_vs_bioreactor <- data.frame(fecal_vs_bioreactor[-30,])

colnames(fecal_vs_bioreactor) <- "family_GM_only"

for (i in 1:length(bioreactor_vs_fecal$family_samples_only))
{
  idx <- which(taxonomyBioreactors$Family == bioreactor_vs_fecal$family_samples_only[i])
  
  tempTax <- taxonomyBioreactors$Phylum[idx]
  
  bioreactor_vs_fecal$phylum_sample_only[i] <- tempTax
}

taxonomyFecalSummary <- taxonomyFecal %>% distinct(phylum, family)

for (i in 1:length(fecal_vs_bioreactor$family_GM_only))
{
  idx <- which(taxonomyFecalSummary$family == fecal_vs_bioreactor$family_GM_only[i])
  
  tempTax <- taxonomyFecalSummary$phylum[idx]
  
  print(tempTax)
  
  fecal_vs_bioreactor$phylum_GM_only[i] <- tempTax
}

write.csv(bioreactor_vs_fecal, "./tables/bioreactor_vs_fecal.csv")

write.csv(as.data.frame(fecal_vs_bioreactor), "./tables/fecal_vs_bioreactor.csv")

```

## Boxplots
```{r}
for (i in bioreactor_vs_fecal$family_samples_only)
{
  p <-
  boxplot_abundance(ps_merge_fam_comp, x = "bioreactor_model", y = i) +
  scale_y_log10() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme(legend.position="bottom")
  
  ggsave(
  paste0("boxplot_", i, ".pdf"),
  p,
  path = "./figures/boxplot",
  width = 25,
  height = 15,
  units = "cm")
}
```

## Pie charts

```{r}
phylum_GM_only <- read.csv("./tables/fecal_vs_bioreactor.csv", row.names = 1)

phylum_GM_only <- phylum_GM_only %>% select(phylum_GM_only) %>% 
  group_by(phylum_GM_only) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

phylum_GM_only <- phylum_GM_only[25:34,]

new <- c("Other", 24, as.numeric(1-sum(phylum_GM_only$perc)), scales::percent(1-sum(phylum_GM_only$perc)))

phylum_GM_only <- rbind(phylum_GM_only,new)

phylum_GM_only$perc <- as.numeric(phylum_GM_only$perc)

phylum_GM_only$n <- as.numeric(phylum_GM_only$n)

# Get the positions
phylum_GM_only2 <- phylum_GM_only %>% 
  mutate(csum = rev(cumsum(rev(perc))), 
         pos = perc/2 + lead(csum, 1),
         pos = if_else(is.na(pos), perc/2, pos))

phylum_samples_only <- read.csv("./tables/bioreactor_vs_fecal.csv", row.names = 1)

phylum_samples_only <- phylum_samples_only %>% select(phylum_sample_only) %>% 
  group_by(phylum_sample_only) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

phylum_samples_only2 <- phylum_samples_only %>% 
  mutate(csum = rev(cumsum(rev(perc))), 
         pos = perc/2 + lead(csum, 1),
         pos = if_else(is.na(pos), perc/2, pos))

# plots

pallete_1 <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F", "#FF7F00", "#CAB2D6", "#6A3D9A", "#FFFF99")

pallete_2 <- c("#CAB2D6", "#FDBF6F", "#B15928", "#d9ef8b", "#808000", "#005a32", "#FF7F00", "#1F78B4", "#FB9A99", "#6A3D9A")


plot_phylum_GM <- ggplot(phylum_GM_only, aes(x = "" , y = perc, fill = fct_inorder(phylum_GM_only))) +
  geom_col(width = 1, color = 1) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = pallete_1)+
  geom_label_repel(data = phylum_GM_only2,
                   aes(y = pos, label = paste0(round(perc*100, digits = 1), "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Phylum")) +
  theme_void()



plot_phylum_samples <- ggplot(phylum_samples_only, aes(x = "" , y = perc, fill = fct_inorder(phylum_samples_only$phylum_sample_only))) +
  geom_col(width = 1, color = 1) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = pallete_2) +
  geom_label_repel(data = phylum_samples_only2,
                   aes(y = pos, label = paste0(round(perc*100, digits = 1), "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Phylum")) +
  theme_void()

arrange_plot <-
  ggarrange(
    plot_phylum_samples,
    plot_phylum_GM,
    ncol = 1,
    nrow = 2,
    labels = c("A.", "B."),
    common.legend = FALSE,
    legend = "bottom"
  )

ggsave(
  "taxonomy_compare.pdf",
  arrange_plot,
  path = "./figures/pie_charts",
  width = 35,
  height = 25,
  units = "cm"
)
```