---
title: "ASVs_enrichment_analysis"
author: "David Garcia"
date: '2023-02-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r results='hide'}
# libraries
pkgs <- c(
  "tidyverse",
  "phyloseq",
  "ggpubr",
  "ggplot2",
  "vegan",
  "reshape2",
  "plotly",
  "ampvis2",
  "microbiomeutilities",
  "RColorBrewer",
  "viridis",
  "biomeUtils",
  "microViz",
  "microbiome",
  "janitor"
)
lapply(pkgs, require, character.only = TRUE)
theme_set(theme_bw())
```

#Create output folders
```{r}
dir.create(path = "./tables/enrichment_analysis/M-ARCOL_Deschamps_et_al_2020", recursive = TRUE)
dir.create(path = "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021")
dir.create(path = "./tables/enrichment_analysis/ARCOL_Leclerc_et_al_2021")
dir.create(path = "./tables/enrichment_analysis/SIMGI_Zorraquin_et_al_2021")
dir.create(path = "./tables/enrichment_analysis/SHIME_Liu_et_al_2022")
```

# M-ARCOL_Deschamps_et_al_2020

## ASVs driven analysis

### Load data

```{r}
load("./RData/primary_id/ps_SRP266568.Rda")

asv_table <- data.frame(ps@otu_table)

asv_table <- t(asv_table) |> data.frame()
```

### Filter data from donor 1
```{r}
# Filter data from donor 1

asv_table_donor_1 <- asv_table %>% select(SRR11963015, SRR11962989, SRR11962990)

# Identify ASVs in faecal sample

asv_table_donor_1_faecal <- asv_table_donor_1 %>%
  filter(SRR11963015 > 0) %>%
  select(SRR11963015)

write.csv(asv_table_donor_1_faecal, "./tables/enrichment_analysis/M-ARCOL_Deschamps_et_al_2020/M-ARCOL_donor_1_faecal.csv")

# Identify the enriched ASVs in lumenal samples

asv_table_donor_1_lumenal <- asv_table_donor_1 %>%
  filter(SRR11963015 == 0 & SRR11962989 > 0) %>%
  select(SRR11962989)

# Identify the enriched ASVs in mucin beads samples

asv_table_donor_1_mucosal <- asv_table_donor_1 %>%
  filter(SRR11963015 == 0 & SRR11962990 > 0) %>%
  select(SRR11962990)
```
### Export to csv files

```{r}
write.csv(asv_table_donor_1_faecal, "./tables/enrichment_analysis/M-ARCOL_Deschamps_et_al_2020/M-ARCOL_donor_1_faecal.csv")

write.csv(asv_table_donor_1_lumenal, "./tables/enrichment_analysis/M-ARCOL_Deschamps_et_al_2020/M-ARCOL_donor_1_lumenal.csv")

write.csv(asv_table_donor_1_mucosal, "./tables/enrichment_analysis/M-ARCOL_Deschamps_et_al_2020/M-ARCOL_donor_1_mucosal.csv")
```

### Filter data from donor 2
```{r}
# Filter_data_from_donor_2

asv_table_donor_2 <- asv_table %>% select(SRR11962962, SRR11963062, SRR11963063)

# Identify ASVs in faecal sample

asv_table_donor_2_faecal <- asv_table_donor_2 %>%
  filter(SRR11962962 > 0) %>%
  select(SRR11962962)

# Identify the enriched ASVs in lumenal samples

asv_table_donor_2_lumenal <- asv_table_donor_2 %>%
  filter(SRR11962962 == 0 & SRR11963062 > 0) %>%
  select(SRR11963062)

# Identify the enriched ASVs in mucin beads samples

asv_table_donor_2_mucosal <- asv_table_donor_2 %>%
  filter(SRR11962962 == 0 & SRR11963063 > 0) %>%
  select(SRR11963063)
```
### Export to csv files

```{r}
write.csv(asv_table_donor_2_faecal, "./tables/enrichment_analysis/M-ARCOL_Deschamps_et_al_2020/M-ARCOL_donor_2_faecal.csv")

write.csv(asv_table_donor_2_lumenal, "./tables/enrichment_analysis/M-ARCOL_Deschamps_et_al_2020/M-ARCOL_donor_2_lumenal.csv")

write.csv(asv_table_donor_2_mucosal, "./tables/enrichment_analysis/M-ARCOL_Deschamps_et_al_2020/M-ARCOL_donor_2_mucosal.csv")
```

# SHIME_Firrman_et_al_2021

## ASVs driven analysis

### Load data
```{r}
load("./RData/primary_id/ps_SRP290868.Rda")

asv_table <- data.frame(ps@otu_table)

asv_table <- t(asv_table) |> data.frame()

selected_samples <- c("SRR12968038", "SRR12968039", "SRR12968045", "SRR12968047", "SRR12968049", "SRR12968051", "SRR12968054", "SRR12968056", "SRR12968043", "SRR12968046", "SRR12968048", "SRR12968050", "SRR12968052", "SRR12968055", "SRR12968084", "SRR12968085", "SRR12968042", "SRR12968087")

asv_table <- asv_table %>% select(all_of(selected_samples))
```

### Analysis by restriction (pick ASVs absent in faecal samples and present in both bioreactor replicates)

### Filter data from single stage bioreactor

```{r}
asv_table_single_stage <- asv_table %>% select(SRR12968087, SRR12968038, SRR12968039, SRR12968084, SRR12968085)

# Identify ASVs in faecal sample

fecal_single_stage <- asv_table_single_stage %>%
  select(SRR12968087) %>%
  filter(SRR12968087 > 0)

# Identify ASV in lumenal samples

lumenal_single <- asv_table_single_stage %>%
  filter(SRR12968087 == 0 & SRR12968038 > 0 & SRR12968039 > 0) %>%
  select(SRR12968038)

mucosal_single <- asv_table_single_stage %>%
  filter(SRR12968087 == 0 & SRR12968084 > 0 & SRR12968085 > 0) %>%
  select(SRR12968084)
```
### Export to csv file

```{r}
write.csv(fecal_single_stage, "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021/SHIME_Firrman_single_stage_faecal.csv")

write.csv(lumenal_single, "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021/SHIME_Firrman_single_lumenal.csv")

write.csv(mucosal_single, "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021/SHIME_Firrman_single_mucosal.csv")
```

### Filter data from multi stage bioreactor

```{r}
asv_table_multi_stage <- asv_table %>% select(SRR12968042, SRR12968045, SRR12968047, SRR12968049, SRR12968051, SRR12968054, SRR12968056, SRR12968048, SRR12968050, SRR12968043, SRR12968046, SRR12968052, SRR12968055)

faecal_multi_stage <- asv_table_multi_stage %>%
  select(SRR12968042) %>%
  filter(SRR12968042 > 0)

lumenal_dc <- asv_table_multi_stage %>%
  filter(SRR12968042 == 0 & SRR12968045 > 0 & SRR12968047 > 0) %>%
  select(SRR12968045)

lumenal_tc <- asv_table_multi_stage %>%
  filter(SRR12968042 == 0 & SRR12968049 > 0 & SRR12968051 > 0) %>%
  select(SRR12968049)

lumenal_ac <- asv_table_multi_stage %>%
  filter(SRR12968042 == 0 & SRR12968054 > 0 & SRR12968056 > 0) %>%
  select(SRR12968054)

mucosal_dc <- asv_table_multi_stage %>%
  filter(SRR12968042 == 0 & SRR12968048 > 0 & SRR12968050 > 0) %>%
  select(SRR12968048)

mucosal_tc <- asv_table_multi_stage %>%
  filter(SRR12968042 == 0 & SRR12968043 > 0 & SRR12968046 > 0) %>%
  select(SRR12968043)

mucosal_ac <- asv_table_multi_stage %>%
  filter(SRR12968042 == 0 & SRR12968052 > 0 & SRR12968055 > 0) %>%
  select(SRR12968052)
```
### Export to csv file

```{r}
write.csv(faecal_multi_stage, "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021/SHIME_Firrman_multi_stage_faecal.csv")

write.csv(lumenal_tc, "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021/SHIME_Firrman_multi_lumenal_tc.csv")

write.csv(lumenal_dc, "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021/SHIME_Firrman_multi_lumenal_dc.csv")

write.csv(lumenal_ac, "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021/SHIME_Firrman_multi_lumenal_ac.csv")

write.csv(mucosal_tc, "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021/SHIME_Firrman_multi_mucosal_tc.csv")

write.csv(mucosal_dc, "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021/SHIME_Firrman_multi_mucosal_dc.csv")

write.csv(mucosal_ac, "./tables/enrichment_analysis/SHIME_Firrman_et_al_2021/SHIME_Firrman_multi_mucosal_ac.csv")
```

# ARCOL_Leclerc_et_al_2021
## ASVs driven analysis

### Load data
```{r}
load("./RData/primary_id/ps_SRP309452.Rda")

asv_table <- data.frame(ps@otu_table)

asv_table <- t(asv_table) |> data.frame()
```
### Filter data from donor 1
```{r}
# Filter data from donor 1

asv_table_donor_1 <- asv_table %>% select(SRR13863485, SRR13863481, SRR13863469)

# Identify ASVs in faecal sample

asv_table_donor_1_faecal <- asv_table_donor_1 %>%
  filter(SRR13863485 > 0) %>%
  select(SRR13863485)

# Identify the enriched ASVs in lumenal samples

asv_table_donor_1_tto_a <- asv_table_donor_1 %>%
  filter(SRR13863485 == 0 & SRR13863481 > 0) %>%
  select(SRR13863481)

# Identify the enriched ASVs in mucin beads samples

asv_table_donor_1_tto_b <- asv_table_donor_1 %>%
  filter(SRR13863485 == 0 & SRR13863469 > 0) %>%
  select(SRR13863469)
```

### Export to csv files

```{r}
write.csv(asv_table_donor_1_faecal, "./tables/enrichment_analysis/ARCOL_Leclerc_et_al_2021/ARCOL_donor_1_faecal.csv")

write.csv(asv_table_donor_1_tto_a, "./tables/enrichment_analysis/ARCOL_Leclerc_et_al_2021/ARCOL_donor_1_tto_a.csv")

write.csv(asv_table_donor_1_tto_b, "./tables/enrichment_analysis/ARCOL_Leclerc_et_al_2021/ARCOL_donor_1_tto_b.csv")
```

### Filter data from donor 2
```{r}
# Filter_data_from_donor_2

asv_table_donor_2 <- asv_table %>% select(SRR13863484, SRR13863457, SRR13863445)

# Identify ASVs in faecal sample

asv_table_donor_2_faecal <- asv_table_donor_2 %>%
  filter(SRR13863484 > 0) %>%
  select(SRR13863484)

# Identify the enriched ASVs in lumenal samples

asv_table_donor_2_tto_a <- asv_table_donor_2 %>%
  filter(SRR13863484 == 0 & SRR13863457 > 0) %>%
  select(SRR13863457)

# Identify the enriched ASVs in mucin beads samples

asv_table_donor_2_tto_b <- asv_table_donor_2 %>%
  filter(SRR13863484 == 0 & SRR13863445 > 0) %>%
  select(SRR13863445)
```

### Export to csv files

```{r}
write.csv(asv_table_donor_2_faecal, "./tables/enrichment_analysis/ARCOL_Leclerc_et_al_2021/ARCOL_donor_2_faecal.csv")

write.csv(asv_table_donor_2_tto_a, "./tables/enrichment_analysis/ARCOL_Leclerc_et_al_2021/ARCOL_donor_2_tto_a.csv")

write.csv(asv_table_donor_2_tto_b, "./tables/enrichment_analysis/ARCOL_Leclerc_et_al_2021/ARCOL_donor_2_tto_b.csv")
```
# SIMGI_Zorraquin_et_al_2021
## ASVs driven analysis

### Load data
```{r}
load("./RData/primary_id/ps_SRP320497.Rda")

asv_table <- data.frame(ps@otu_table)

asv_table <- t(asv_table) |> data.frame()
```

### Filter data from control
```{r}
# Filter data from donor 1

asv_table_control <- asv_table %>% select(SRR14585829, SRR14585827)

# Identify ASVs in faecal sample

asv_table_control_faecal <- asv_table_control %>%
  filter(SRR14585829 > 0) %>%
  select(SRR14585829)

# Identify ASVs after fermentation with colon nutrient medium

asv_table_control_48h <- asv_table_control %>%
  filter(SRR14585829 == 0 & SRR14585827 > 0) %>%
  select(SRR14585827)
```

### Export to csv files

```{r}
write.csv(asv_table_control_faecal, "./tables/enrichment_analysis/SIMGI_Zorraquin_et_al_2021/SIMGI_control_faecal.csv")

write.csv(asv_table_control_48h, "./tables/enrichment_analysis/SIMGI_Zorraquin_et_al_2021/SIMGI_control_48h.csv")
```

### Filter data from treated samples (Digested wine - DW)
```{r}
# Filter_data_from_faecal sample and DW

asv_table_DW <- asv_table %>% select(SRR14585826, SRR14585824)

# Identify ASVs after fermentation with colon nutrient medium

asv_table_DW_faecal <- asv_table_DW %>%
  filter(SRR14585826 > 0) %>%
  select(SRR14585826)

# Identify ASVs after fermentation with DW

asv_table_DW_48h <- asv_table_DW %>%
  filter(SRR14585826 == 0 & SRR14585824 > 0) %>%
  select(SRR14585824)
```

### Export to csv files

```{r}
write.csv(asv_table_DW_faecal, "./tables/enrichment_analysis/SIMGI_Zorraquin_et_al_2021/SIMGI_DW_faecal.csv")

write.csv(asv_table_DW_48h, "./tables/enrichment_analysis/SIMGI_Zorraquin_et_al_2021/SIMGI_DW_48h.csv")
```

# SHIME_Liu_et_al_2022
## ASVs driven analysis
### Load data
```{r}
load("./RData/primary_id/ps_SRP359931.Rda")

asv_table <- data.frame(ps@otu_table)

asv_table <- t(asv_table) |> data.frame()

selected_samples <- c(
  "SRR18036853", "SRR18036849", "SRR18036844", "SRR18036839",
  "SRR18036845", "SRR18036840", "SRR18036830", "SRR18036827",
  "SRR18036836", "SRR18036855", "SRR18036842", "SRR18036838",
  "SRR18036831", "SRR18036828", "SRR18036829", "SRR18036826",
  "SRR18036852", "SRR18036848", "SRR18036841", "SRR18036837"
)

asv_table <- asv_table %>% select(all_of(selected_samples))

# Donor A
# Control (SRR18036853) Control_replicate (SRR18036849) Control 48h (SRR18036844) Control 48h replicate (SRR18036839)

# Donor B
# Control (SRR18036845) Control_replicate (SRR18036840) Control 48h (SRR18036830) Control 48h replicate (SRR18036827)

# Donor C
# Control (SRR18036836) Control_replicate (SRR18036855) Control 48h (SRR18036842) Control 48h replicate (SRR18036838)

# Donor D
# Control (SRR18036831) Control_replicate (SRR18036828) Control 48h (SRR18036829) Control 48h replicate (SRR18036826)

# Donor F
# Control (SRR18036852) Control_replicate (SRR18036848) Control 48h (SRR18036841) Control 48h replicate (SRR18036837)
```
### Analysis by restriction
### Filter data from control samples and all donors

```{r}
donor_a_faecal <- asv_table %>%
  select(SRR18036853, SRR18036849) %>%
  filter(SRR18036853 > 0 & SRR18036849 > 0) %>%
  select(SRR18036853)

donor_a_enriched <- asv_table %>%
  select(SRR18036853, SRR18036849, SRR18036844, SRR18036839) %>%
  filter(SRR18036853 == 0 & SRR18036849 == 0 & SRR18036844 > 0 & SRR18036839 > 0) %>%
  select(SRR18036839)

donor_b_faecal <- asv_table %>%
  select(SRR18036845, SRR18036840) %>%
  filter(SRR18036845 > 0 & SRR18036840 > 0) %>%
  select(SRR18036845)

donor_b_enriched <- asv_table %>%
  select(SRR18036845, SRR18036840, SRR18036830, SRR18036827) %>%
  filter(SRR18036845 == 0 & SRR18036840 == 0 & SRR18036830 > 0 & SRR18036827 > 0) %>%
  select(SRR18036827)

donor_c_faecal <- asv_table %>%
  select(SRR18036836, SRR18036855) %>%
  filter(SRR18036836 > 0 & SRR18036855 > 0) %>%
  select(SRR18036836)

donor_c_enriched <- asv_table %>%
  select(SRR18036836, SRR18036855, SRR18036842, SRR18036838) %>%
  filter(SRR18036836 == 0 & SRR18036855 == 0 & SRR18036842 > 0 & SRR18036838 > 0) %>%
  select(SRR18036838)

donor_d_faecal <- asv_table %>%
  select(SRR18036831, SRR18036828) %>%
  filter(SRR18036831 > 0 & SRR18036828 > 0) %>%
  select(SRR18036831)

donor_d_enriched <- asv_table %>%
  select(SRR18036831, SRR18036828, SRR18036829, SRR18036826) %>%
  filter(SRR18036831 == 0 & SRR18036828 == 0 & SRR18036829 > 0 & SRR18036826 > 0) %>%
  select(SRR18036826)

donor_f_faecal <- asv_table %>%
  select(SRR18036852, SRR18036848) %>%
  filter(SRR18036852 > 0 & SRR18036848 > 0) %>%
  select(SRR18036852)

donor_f_enriched <- asv_table %>%
  select(SRR18036852, SRR18036848, SRR18036841, SRR18036837) %>%
  filter(SRR18036852 == 0 & SRR18036848 == 0 & SRR18036841 > 0 & SRR18036837 > 0) %>%
  select(SRR18036837)
```
### Export to csv

```{r}
write.csv(donor_a_faecal, "./tables/enrichment_analysis/SHIME_Liu_et_al_2022/SHIME_donor_a_faecal.csv")

write.csv(donor_a_enriched, "./tables/enrichment_analysis/SHIME_Liu_et_al_2022/SHIME_donor_a_enriched.csv")

write.csv(donor_b_faecal, "./tables/enrichment_analysis/SHIME_Liu_et_al_2022/SHIME_donor_b_faecal.csv")

write.csv(donor_b_enriched, "./tables/enrichment_analysis/SHIME_Liu_et_al_2022/SHIME_donor_b_enriched.csv")

write.csv(donor_c_faecal, "./tables/enrichment_analysis/SHIME_Liu_et_al_2022/SHIME_donor_c_faecal.csv")

write.csv(donor_c_enriched, "./tables/enrichment_analysis/SHIME_Liu_et_al_2022/SHIME_donor_c_enriched.csv")

write.csv(donor_d_faecal, "./tables/enrichment_analysis/SHIME_Liu_et_al_2022/SHIME_donor_d_faecal.csv")

write.csv(donor_d_enriched, "./tables/enrichment_analysis/SHIME_Liu_et_al_2022/SHIME_donor_d_enriched.csv")

write.csv(donor_f_faecal, "./tables/enrichment_analysis/SHIME_Liu_et_al_2022/SHIME_donor_f_faecal.csv")

write.csv(donor_f_enriched, "./tables/enrichment_analysis/SHIME_Liu_et_al_2022/SHIME_donor_f_enriched.csv")
```
